```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(data.table)
library(ggplot2)
library(igraph)
library(reshape2)
library(tidyverse)
library(scales)
library(stringr)
library(psych)  #corr.test {psych}
theme_set(theme_light())
```



```{r}
make_pois_community <- function(num_samples = 100, num_otus = 10, lambda = 1, names = "OTU",
                           sites = "Sample", fun = f) {
  matrix(rpois(num_samples * num_otus, lambda),
                        nrow = num_otus, ncol = num_samples,
                        dimnames = list(
                          paste(names, 1:num_otus, sep = "_"),
                          paste(sites, 1:num_samples, sep = "_")
                        )
  
)
}

make_unif_community<- function(num_samples = 50, num_otus = 10, lower = 100, upper = 120, names = "OTU",
                           sites = "Sample", fun = f) {
  matrix(runif(num_samples * num_otus, lower, upper),
                        nrow = num_otus, ncol = num_samples,
                        dimnames = list(
                          paste(names, 1:num_otus, sep = "_"),
                          paste(sites, 1:num_samples, sep = "_")
                        )
  )
}



make_negative_community<- function(num_samples = 50, num_otus = 10, lambda = 4, names = "OTU",
                            sites = "Sample", fun = f){
  matrix(1000 - rpois(num_samples * num_otus, lambda),
                        nrow = num_otus, ncol = num_samples,
                        dimnames = list(
                          paste(names, 1:num_otus, sep = "_"),
                          paste(sites, 1:num_samples, sep = "_")
                        )
  )
}

make_positive_community <- function(num_samples = 50, num_otus = 10, lambda = 4, names = "OTU",
                            sites = "Sample", fun = f){
  matrix(5 * rpois(num_samples * num_otus, lambda),
                        nrow = num_otus, ncol = num_samples,
                        dimnames = list(
                          paste(names, 1:num_otus, sep = "_"),
                          paste(sites, 1:num_samples, sep = "_")
                        )
  )
}
  

```


```{r}
comm_1 <- make_pois_community(names = "COMM_A", num_samples = 100, num_otus = 5, lambda = 100)
comm_2 <- make_negative_community(names = "COMM_C", num_samples = 100, num_otus = 5, lambda = 100)
comm_3 <- make_positive_community(names = "COMM_D", num_samples = 100, num_otus = 5, lambda = 4000)
noise <- make_pois_community(num_otus = 50)
sim_community <- rbind(comm_1, comm_2, comm_3, noise)
```

#plot_dd function can plot degree distribution of graph object
```{r plot_dd.function}
plot_dd <- function(g){
  g.dd <- data.frame(0:(length(degree_distribution(g)) - 1), 
                     degree_distribution(g, cumulative = FALSE))
  colnames(g.dd) <-c("k", "p_k")
  ggplot(g.dd, aes(x = k, y = p_k)) + geom_bar(stat = "identity") +
    labs(x = "Degree (k)", y = "Proportion of nodes of degree k (p(k))", title = "Degree distribution: k vs. p(k)", 
         subtitle = paste("Graph:", deparse(substitute(g)))) +
    theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) + 
    scale_x_continuous(breaks = function(x) seq(ceiling(x[1]), floor(x[2]), by = 1))
}
```

#getting correlation matrix (OTU BY OTU)
```{r}
sim_community.cor_matrix <- cor(t(sim_community), method = "spearman")
```


#getting correlation matrix with different correction method by changing the adjust = "" in the first line of code
p.adjust.methods
# ("holm", "hochberg", "hommel", "bonferroni", "BH", "BY", "fdr", "none")
```{r test case}
test_network <- corr.test(sim_community.cor_matrix, y = NULL, use = "pairwise", method="spearman", adjust = "BH",alpha=0.05) 
a <- test_network$r[-1]
b <- test_network$p
test_sig_correlation =ifelse( b < 0.05, a, NA)
cor.plot(test_sig_correlation)
threshold <-test_sig_correlation
threshold[threshold >= 0.3] <- 1
threshold[threshold <= -0.3] <- -1
threshold[threshold >= -0.3 & threshold <= 0.3] <- 0
threshold_graph <- graph_from_adjacency_matrix(threshold, mode ="undirected", diag =FALSE)
threshold_graph.copy <- delete.vertices(threshold_graph, degree(threshold_graph) == 0)
plot(threshold_graph.copy, 
     layout = layout.fruchterman.reingold(threshold_graph.copy),
     main = "Network with isolated nodes removed")
cfg <- cluster_fast_greedy(threshold_graph.copy)
plot(cfg, threshold_graph.copy)
ceb<- cluster_edge_betweenness(threshold_graph.copy)
plot(ceb,
     threshold_graph.copy)
E(threshold_graph.copy) #check out the edge attribute of the graph
which.max(degree(threshold_graph.copy)) #check out the highest degree nodes
plot_dd(threshold_graph)
plot_dd(threshold_graph.copy)
```



STEP BY STEP:
#1. selecting only significant OTU Pairs by thresholding p-value <0.05 and correlation values not equal to 1
```{r}
a <- test_network$r[-1] #dropping correlation values that are 1
b <- test_network$p
sig_correlation =ifelse( b < 0.05, a, NA)
```

#2. getting correlation plot(heatmap)
```{r}
cor.plot(sig_correlation)
```

#3. thresholding the values of this correlation matrix. We say that two OTUs are *positively correlated* if the Spearman correlation between them is grater than $0.30$.
```{r}
threshold <- sig_correlation
threshold[threshold >= 0.3] <- 1 #Assigning 1 to all the values of the matrix above our cutoff:
threshold[threshold <= -0.3] <- -1 #Doing the same for negatively co
threshold[threshold >= -0.3 & threshold <= 0.3] <- 0 #Setting everything else to 0 since they don't fall into our definition                                                        of positive or negatively correlated
```

#4. make correlation graph from threshold correlation matrix
```{r}
threshold_graph <- graph_from_adjacency_matrix(threshold, mode ="undirected", diag =FALSE)
```

#5. plot correlation graph
```{r}
gorder(threshold_graph) #get the number of edges for the graph
plot(threshold_graph, 
     layout = layout.fruchterman.reingold(threshold_graph),
     main = "Network")
```

#6. plot the correlation graoh with isolated nodes removed
```{r}
threshold_graph.copy <- delete.vertices(threshold_graph, degree(threshold_graph) == 0)
gorder(threshold_graph.copy) 
plot(threshold_graph.copy, 
     layout = layout.fruchterman.reingold(threshold_graph.copy),
     main = "Network with isolated nodes removed")
```

#7. plot the degree distribution using plot_dd function
```{r}
plot_dd(threshold_graph.copy)
```

#8. clustering by cluster fast greedy method
```{r}
cfg <- cluster_fast_greedy(threshold_graph.copy)
plot(cfg, threshold_graph.copy)

```

#9. clustering by edge betweeness method
```{r}
ceb <- cluster_edge_betweenness(threshold_graph.copy)
plot(ceb, threshold_graph.copy)
```
